`flippant` --- An `R` package for the automated analysis of dithionite scramblase

### 2016-07-22 inst/manuscript/manuscript.rmd
Magrittrize the microbenchmarking read-out and switch to use of MAD;

@@

# Availablility and Implementation
`flippant` has been implemented in `R` [@r_core_team_r:_2016] and integrated 
into the `Bioconductor` collection of tools for biomedical data analysis [@huber_orchestrating_2015]. It is thus install-- and loadable as follows from 
within a functional `R` environment:

@@

`flippant`'s capabilities are demonstrated here by reanalyzing a subset of the
data underlying Figure 4C--F in @ploier_dimerization_2016. The first step is to
extract the data files, stored in a ZIP archive within the pacakge.

```{r extract_case_study_dataset}
analysis_dir <- file.path(tempdir(), "flippant-case-study")
extract_case_study_data(analysis_dir)
template_file <- file.path(analysis_dir, "inputTable.txt")
```

To begin the analysis it is useful to plot the spectral traces of each 
experimental series (Figure 1). Note
that algorithmic alignment to the time of dithionite addition is employed
(default) as well as  custom trimming of the time axis to account for divergent
acquisition  times between experiments. PPR scaling is used with the default
factor of 0.65.

```{r figure1_creation, fig.height = 4, fig.cap = "**`flippant`--based plotting of spectral traces underlying a subset of the data in Figure 4C--F from @ploier_dimerization_2016.** See text for command and details."}
# Plot spectral traces (Figure 1)
scramblase_assay_traces(
  template_file,
  time_max_sec = 350)
```

@@

```{r noeval_demonstrate_plot, eval = FALSE}
# Plot PPR plot (Figure 2A)
scramblase_assay_plot(template_file)
```

```{r figure2_creation, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "**`flippant`--based reanalysis of a subset of the data in Figure 4C--F from @ploier_dimerization_2016.** (A) and (B) show protein to phospholipid ration (PPR) plots for the data using a x--axis trimmed to a maximum value of 3 (see text for details). In (A) data is separately plotted for independent experiments, while (B) combines all data points, aiming for reliability of fit. 'adj. PPR' indicates that the measure has been scaled to account for a vesicle pool refractory to flippase reconsitution."}
figure2 %<>% multipanelfigure::add_panel(
  top_panel = 1,
  panel = flippant::scramblase_assay_plot(template_file) + 
    ggplot2::theme_bw() + 
    ggplot2::xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ggplot2::ylab(expression(P[{phantom()>=1}])))
figure2 %<>% multipanelfigure::add_panel(
  top_panel = 2,
  panel = flippant::scramblase_assay_plot(template_file, split_by_experiment = FALSE) + 
    ggplot2::theme_bw() + 
    ggplot2::xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ggplot2::ylab(expression(P[{phantom()>=1}])) +
    ggplot2::facet_wrap(c("`Experimental Series`"))+
    ggplot2::guides(colour = "none"))
figure2
```

@@

```{r noeval_demonstrate_plot2, eval = FALSE}
# Plot PPR plot (Figure 2B)
scramblase_assay_plot(
  template_file,
  split_by_experiment = FALSE)
```

@@

```{r noeval_demonstrate_tabular_writeout, eval = FALSE}
# Write out results in tabular manner (Table 1)
scramblase_assay_stats(
  template_file,
  split_by_experiment = FALSE)
```

@@

analysis_timing_median <- signif(
  summary(
    analysis_timing)$median,
  2)

analysis_timing_iqr <- signif(
  diff(
    unlist(
      summary(analysis_timing)[c("lq", "uq")],
      use.names = FALSE)),
  2)
```

With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
R as an environment for statistical computing and graphics 
[@r_core_team_r:_2016], scientists can quickly analyse such data and arrive at 
publication--grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
`R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr`~\text{s}$; $n=`r analysis_timing_n`$).

>>>>

tmpDir <- tempdir()
file.path("extdata", "PloierEtAl_Data.zip") %>%
  system.file(
    package = "flippant",
    mustWork = TRUE) %>%
  unzip(
    overwrite = TRUE,
    exdir = tmpDir)

@@

# Availablility and Implementation
`flippant` has been implemented in `R` [@r_core_team_r:_2016] and integrated 
into the `bioconductor` collection of tools for biomedical data analysis [@huber_orchestrating_2015]. It is thus install-- and loadable as follows from 
within a functional `R` environment:

@@

`flippant`s capabilities are demonstrated here by reanalyzing a subset of the
data underlying  Figure 4C--F in @ploier_dimerization_2016. The following work
flow in a first  step plots the corresponding spectral traces (Figure 1). Note
that algorithmic alignment to the time of dithionite addition is employed
(default) as well as  custom trimming of the time axis to account for divergent
acquisition  times between experiments. PPR scaling is used with the default
factor of 0.65.

```{r noeval_demonstrate_traces, eval = FALSE}
# Plot spectral traces (Figure 1)
scramblase_assay_traces(
  "inputTable.txt",
  time_max_sec = 350)
```

```{r figure1_creation, echo = FALSE, fig.height = 4, fig.cap = "**`flippant`--based plotting of spectral traces underlying a subset of the data in Figure 4C--F from @ploier_dimerization_2016.** See text for command and details."}
# Plot spectral traces (Figure 1)
x <- tmpDir %>%
  file.path("inputTable.txt") %>%
  flippant:::read_scramblase_input_file()
x[["Path"]] <- file.path(tmpDir, x[["Path"]])
flippant::scramblase_assay_traces(
  x,
  time_max_sec = 350)
```

@@

```{r noeval_demonstrate_plot, eval = FALSE}
# Plot PPR plot (Figure 2A)
scramblase_assay_plot("inputTable.txt")
```

```{r figure2_creation, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "**`flippant`--based reanalysis of a subset of the data in Figure 4C--F from @ploier_dimerization_2016.** (A) and (B) show protein to phospholipid ration (PPR) plots for the data using a x--axis trimmed to a maximum value of 3 (see text for details). In (A) data is separately plotted for independent experiments, while (B) combines all data points, aiming for reliability of fit. 'adj. PPR' indicates that the measure has been scaled to account for a vesicle pool refractory to flippase reconsitution."}
figure2 %<>% multipanelfigure::add_panel(
  top_panel = 1,
  panel = flippant::scramblase_assay_plot(x) + 
    ggplot2::theme_bw() + 
    ggplot2::xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ggplot2::ylab(expression(P[{phantom()>=1}])))
figure2 %<>% multipanelfigure::add_panel(
  top_panel = 2,
  panel = flippant::scramblase_assay_plot(x, split_by_experiment = FALSE) + 
    ggplot2::theme_bw() + 
    ggplot2::xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ggplot2::ylab(expression(P[{phantom()>=1}])) +
    ggplot2::facet_wrap(c("`Experimental Series`"))+
    ggplot2::guides(colour = "none"))
figure2
```

@@

```{r noeval_demonstrate_plot2, eval = FALSE}
# Plot PPR plot (Figure 2B)
scramblase_assay_plot(
  "inputTable.txt",
  split_by_experiment = FALSE)
```

@@

```{r noeval_demonstrate_tabular_writeout, eval = FALSE}
# Write out results in tabular manner (Table 1)
scramblase_assay_stats(
  "inputTable.txt",
  split_by_experiment = FALSE)
```

@@

analysis_timing_median <- analysis_timing %>%
  summary() %>%
  magrittr::extract2("median") %>%
  signif(2)

analysis_timing_mad <- analysis_timing %>%
  magrittr::extract2("time") %>%
  mad() %>%
  magrittr::divide_by(1E9) %>%
  signif(2)
```

With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
R as an environment for statistical computing and graphics 
[@r_core_team_r:_2016], scientists can quickly analyse such data and arrive at 
publication--grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
`R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{MAD}=`r analysis_timing_mad`~\text{s}$; $n=`r analysis_timing_n`$).
